---
title: "WF_NaturalMortality_Analyses"
author: "Kamran Walsh"
date: "2026-02-10"
output: html_document
---

Code for running analyses for WF natural mortality project

Load packages
```{r}
library(dplyr)
library(tidyr)
library(data.table)
library(mgcv)
library(purrr)
library(GGally)
library(tibble)
library(mgcv)
library(gratia)
```

#CATCH-AT-AGE COMPOSITIONS w/PLUS GROUP 10+ 

#Index catch-at-age code for GB 
```{r}
#the code for plotting the age comps is from the WHAM source code 

bubble.col = "white"

GB_agecomps <- read.csv("~/Desktop/StockAssessmentPracticum/WF_NaturalMortality/STOCKEFF_SV_172905_GBK_NONE_strat_mean_age.csv")

#turn file into asap matrix format 
GB_agecomps_filter <- na.omit(GB_agecomps)

GB_agecomps_matrix <- GB_agecomps_filter %>% 
select(STOCK_ABBREV, SEASON, YEAR, AGE, NO_AT_AGE) %>%
  pivot_wider(
    names_from = AGE,
    values_from = NO_AT_AGE
  )
col_nums <- sort(as.numeric(names(GB_agecomps_matrix[4:24])))
GB_agecomps_matrix <- GB_agecomps_matrix %>%
  select(STOCK_ABBREV, SEASON, YEAR, all_of(as.character(col_nums)))

#PLOT AGE COMPS BY SEASON 

#GB spring 

#arrange data 
GB_agecomps_matrix_spring <- GB_agecomps_matrix %>% filter(SEASON == "SPRING") %>%
  mutate(across(everything(), ~replace(., is.na(.), 0)))
years = GB_agecomps_matrix_spring$YEAR
nyrs = length(years)
ages = c(0:10)
ages.lab = c(as.character(ages[1:10]), "10+")
n_ages = length(ages)

#calculate proportion-at-age 
#acomp.obs = GB_agecomps_matrix_spring %>% select(4:24)

acomp.obs = GB_agecomps_matrix_spring %>% mutate('10' = rowSums(GB_agecomps_matrix_spring[, 14:24], na.rm = TRUE)) %>% select(4:14)
names(acomp.obs) <- as.numeric(ages)
catch.yrs = nrow(GB_agecomps_matrix_spring)

#plot
my.title <- paste0("Catch-at-Age for GB NMFS Spring Survey")
  origpar <- par(no.readonly = TRUE)
  par(mar=c(4,4,2,2), oma=c(1,1,1,1), mfrow=c(1,1))
  scale.catch.obs <- 5
  z3 <- as.matrix(acomp.obs)# * scale.catch.obs

  plot(ages, rev(ages),  xlim = range(ages), ylim = c(years[nyrs],(years[1]-2)), xlab = "Age", ylab = "", type = "n", axes=FALSE)
  axis(1, at= ages, lab = ages.lab) #change this depending on plus group 
  axis(2, at = rev(years), lab = rev(years), cex.axis=0.75, las=1)
  box()
  abline(h=years, col="lightgrey")
  segments(x0=ages, y0=rep(years[1],n_ages), x1=ages, y1=rep(years[nyrs],n_ages), col = "lightgrey", lty = 1)
  if (length(catch.yrs)>0) for (j in 1:nyrs) points(ages, rep(years[j], n_ages), cex=z3[j,], col="black",# bg = bubble.col,
                                                    pch = 21)

  bubble.legend1 <- c(0.05,0.2,0.4)
  bubble.legend2 <- bubble.legend1 * scale.catch.obs
  legend("topright", xpd=TRUE, legend=bubble.legend1, pch=rep(21, 3), pt.cex=bubble.legend2, horiz=T , col='black', pt.bg = bubble.col)
  title (my.title, outer=TRUE, line=-1)
  par(origpar)
  
#GB Fall
#arrange data 
GB_agecomps_matrix_fall <- GB_agecomps_matrix %>% filter(SEASON == "FALL") %>%
  mutate(across(everything(), ~replace(., is.na(.), 0)))
years = GB_agecomps_matrix_fall$YEAR
nyrs = length(years)
ages = c(0:10)
ages.lab = c(as.character(ages[1:10]), "10+")
n_ages = length(ages)

#calculate proportion-at-age 
#acomp.obs = GB_agecomps_matrix_fall %>% select(4:24)

acomp.obs = GB_agecomps_matrix_fall %>% mutate('10' = rowSums(GB_agecomps_matrix_fall[, 14:24], na.rm = TRUE)) %>% select(4:14)
names(acomp.obs) <- as.numeric(ages)
catch.yrs = nrow(GB_agecomps_matrix_fall)

#plot
my.title <- paste0("Catch-at-Age for GB NMFS Fall Survey")
  origpar <- par(no.readonly = TRUE)
  par(mar=c(4,4,2,2), oma=c(1,1,1,1), mfrow=c(1,1))
  scale.catch.obs <- 5
  z3 <- as.matrix(acomp.obs) #* scale.catch.obs

  plot(ages, rev(ages),  xlim = range(ages), ylim = c(years[nyrs],(years[1]-2)), xlab = "Age", ylab = "", type = "n", axes=FALSE)
  axis(1, at= ages, lab = ages.lab)
  axis(2, at = rev(years), lab = rev(years), cex.axis=0.75, las=1)
  box()
  abline(h=years, col="lightgrey")
  segments(x0=ages, y0=rep(years[1],n_ages), x1=ages, y1=rep(years[nyrs],n_ages), col = "lightgrey", lty = 1)
  if (length(catch.yrs)>0) for (j in 1:nyrs) points(ages, rep(years[j], n_ages), cex=z3[j,], col="black",# bg = bubble.col,
                                                    pch = 21)

  bubble.legend1 <- c(0.05,0.2,0.4)
  bubble.legend2 <- bubble.legend1 * scale.catch.obs
  legend("topright", xpd=TRUE, legend=bubble.legend1, pch=rep(21, 3), pt.cex=bubble.legend2, horiz=T , col='black', pt.bg = bubble.col)
  title (my.title, outer=TRUE, line=-1)
  par(origpar)
```

#Index catch-at-age code for GOM
```{r}
bubble.col = "white"

GOM_agecomps <- read.csv("~/Desktop/StockAssessmentPracticum/WF_NaturalMortality/STOCKEFF_SV_172905_GOM_NONE_strat_mean_age.csv")

#different surveys -> NMFS BTS + MADMF

#NMFS BTS

#turn file into asap matrix format 
GOM_agecomps_filter <- na.omit(GOM_agecomps)

GOM_agecomps_matrix_nmfs <- GOM_agecomps_filter %>%
  filter(grepl("NMFS", SURVEY)) %>%
select(STOCK_ABBREV, SEASON, YEAR, AGE, NO_AT_AGE) %>%
  pivot_wider(
    names_from = AGE,
    values_from = NO_AT_AGE
  )
col_nums <- sort(as.numeric(names(GOM_agecomps_matrix_nmfs[4:19])))
GOM_agecomps_matrix_nmfs <- GOM_agecomps_matrix_nmfs %>%
  select(STOCK_ABBREV, SEASON, YEAR, all_of(as.character(col_nums))) 

#PLOT AGE COMPS BY SEASON 

#GOM spring 

#arrange data 
GOM_agecomps_matrix_nmfs_spring <- GOM_agecomps_matrix_nmfs %>% filter(SEASON == "SPRING") %>%
  mutate(across(everything(), ~replace(., is.na(.), 0)))
years = GOM_agecomps_matrix_nmfs_spring$YEAR
nyrs = length(years)
ages = c(0:7)
ages.lab = c(as.character(ages[1:7]), "7+")
n_ages = length(ages)

#calculate proportion-at-age 
#acomp.obs = GOM_agecomps_matrix_nmfs_spring %>% select(4:19)

acomp.obs = GOM_agecomps_matrix_nmfs_spring %>% mutate('7' = rowSums(GOM_agecomps_matrix_nmfs_spring[, 11:19], na.rm = TRUE)) %>% select(4:11)
names(acomp.obs) <- as.numeric(ages)
catch.yrs = nrow(GOM_agecomps_matrix_nmfs_spring)

#plot
my.title <- paste0("NEFSC Bottom Trawl Survey GOM Spring Abundance-at-Age")
  origpar <- par(no.readonly = TRUE)
  par(mar=c(4,4,2,2), oma=c(1,1,1,1), mfrow=c(1,1))
  scale.catch.obs <- 5
  z3 <- as.matrix(acomp.obs) # * scale.catch.obs

  plot(ages, rev(ages),  xlim = range(ages), ylim = c(years[nyrs],(years[1]-2)), xlab = "Age", ylab = "", type = "n", axes=FALSE)
  axis(1, at= ages, lab = ages.lab)
  axis(2, at = rev(years), lab = rev(years), cex.axis=0.75, las=1)
  box()
  abline(h=years, col="lightgray")
  segments(x0=ages, y0=rep(years[1],n_ages), x1=ages, y1=rep(years[nyrs],n_ages), col = "lightgray", lty = 1)
  if (length(catch.yrs)>0) for (j in 1:nyrs) points(ages, rep(years[j], n_ages), cex=z3[j,], col="black", #bg = bubble.col,
                                                    pch = 21)

  bubble.legend1 <- c(0.05,0.2,0.4)
  bubble.legend2 <- bubble.legend1 * scale.catch.obs
  legend("topright", xpd=TRUE, legend=bubble.legend1, pch=rep(21, 3), pt.cex=bubble.legend2, horiz=T , col='black', pt.bg = bubble.col)
  title (my.title, outer=TRUE, line=-1)
  par(origpar)
  
#GOM Fall
  
#arrange data 
GOM_agecomps_matrix_nmfs_fall <- GOM_agecomps_matrix_nmfs %>% filter(SEASON == "FALL") %>%
  mutate(across(everything(), ~replace(., is.na(.), 0)))
years = GOM_agecomps_matrix_nmfs_fall$YEAR
nyrs = length(years)
ages = c(0:7)
ages.lab = c(as.character(ages[1:7]), "7+")
n_ages = length(ages)

#calculate proportion-at-age 
#acomp.obs = GOM_agecomps_matrix_nmfs_fall %>% select(4:19)

acomp.obs = GOM_agecomps_matrix_nmfs_fall %>% mutate('7' = rowSums(GOM_agecomps_matrix_nmfs_fall[, 11:19], na.rm = TRUE)) %>% select(4:11)
names(acomp.obs) <- as.numeric(ages)
catch.yrs = nrow(GOM_agecomps_matrix_nmfs_fall)

#plot
my.title <- paste0("NEFSC Bottom Trawl Survey GOM Fall Abundance-at-Age")
  origpar <- par(no.readonly = TRUE)
  par(mar=c(4,4,2,2), oma=c(1,1,1,1), mfrow=c(1,1))
  scale.catch.obs <- 5
  z3 <- as.matrix(acomp.obs) #* scale.catch.obs

  plot(ages, rev(ages),  xlim = range(ages), ylim = c(years[nyrs],(years[1]-2)), xlab = "Age", ylab = "", type = "n", axes=FALSE)
  axis(1, at= ages, lab = ages.lab)
  axis(2, at = rev(years), lab = rev(years), cex.axis=0.75, las=1)
  box()
  abline(h=years, col="lightgray")
  segments(x0=ages, y0=rep(years[1],n_ages), x1=ages, y1=rep(years[nyrs],n_ages), col = "lightgray", lty = 1)
  if (length(catch.yrs)>0) for (j in 1:nyrs) points(ages, rep(years[j], n_ages), cex=z3[j,], col="black", #bg = bubble.col,
                                                    pch = 21)

  bubble.legend1 <- c(0.05,0.2,0.4)
  bubble.legend2 <- bubble.legend1 * scale.catch.obs
  legend("topright", xpd=TRUE, legend=bubble.legend1, pch=rep(21, 3), pt.cex=bubble.legend2, horiz=T , col='black', pt.bg = bubble.col)
  title (my.title, outer=TRUE, line=-1)
  par(origpar)

#MADMF -> no need to make plus group correction since this isn't going in paper 

#turn file into asap matrix format 
GOM_agecomps_filter <- na.omit(GOM_agecomps)

GOM_agecomps_matrix_madmf <- GOM_agecomps_filter %>%
  filter(grepl("MADMF", SURVEY)) %>%
select(STOCK_ABBREV, SEASON, YEAR, AGE, NO_AT_AGE) %>%
  pivot_wider(
    names_from = AGE,
    values_from = NO_AT_AGE
  )
col_nums <- sort(as.numeric(names(GOM_agecomps_matrix_madmf[4:19])))
GOM_agecomps_matrix_madmf <- GOM_agecomps_matrix_madmf %>%
  select(STOCK_ABBREV, SEASON, YEAR, all_of(as.character(col_nums)))

#PLOT AGE COMPS BY SEASON 

#GOM spring 

#arrange data 
GOM_agecomps_matrix_madmf_spring <- GOM_agecomps_matrix_madmf %>% filter(SEASON == "SPRING") %>%
  mutate(across(everything(), ~replace(., is.na(.), 0)))
years = GOM_agecomps_matrix_madmf_spring$YEAR
nyrs = length(years)
ages = c(0:15)
ages.lab = as.character(ages)
n_ages = length(ages)

#calculate proportion-at-age 
acomp.obs = GOM_agecomps_matrix_madmf_spring %>% select(4:19)
names(acomp.obs) <- as.numeric(ages)
catch.yrs = nrow(GOM_agecomps_matrix_madmf_spring)

#plot
my.title <- paste0("Catch-at-Age for GOM MADMF Spring Survey")
  origpar <- par(no.readonly = TRUE)
  par(mar=c(4,4,2,2), oma=c(1,1,1,1), mfrow=c(1,1))
  scale.catch.obs <- 0.1
  z3 <- as.matrix(acomp.obs)  * scale.catch.obs

  plot(ages, rev(ages),  xlim = range(ages), ylim = c(years[nyrs],(years[1]-2)), xlab = "Age", ylab = "", type = "n", axes=FALSE)
  axis(1, at= ages, lab = ages.lab)
  axis(2, at = rev(years), lab = rev(years), cex.axis=0.75, las=1)
  box()
  abline(h=years, col="lightgray")
  segments(x0=ages, y0=rep(years[1],n_ages), x1=ages, y1=rep(years[nyrs],n_ages), col = "lightgray", lty = 1)
  if (length(catch.yrs)>0) for (j in 1:nyrs) points(ages, rep(years[j], n_ages), cex=z3[j,], col="black", #bg = bubble.col,
                                                    pch = 21)

  bubble.legend1 <- c(0.05,0.2,0.4)
  bubble.legend2 <- bubble.legend1 * scale.catch.obs
  legend("topright", xpd=TRUE, legend=bubble.legend1, pch=rep(21, 3), pt.cex=bubble.legend2, horiz=T , col='black', pt.bg = bubble.col)
  title (my.title, outer=TRUE, line=-1)
  par(origpar) 
  
#GOM Fall
  
#arrange data 
GOM_agecomps_matrix_madmf_fall <- GOM_agecomps_matrix_madmf %>% filter(SEASON == "FALL") %>%
  mutate(across(everything(), ~replace(., is.na(.), 0)))
years = GOM_agecomps_matrix_madmf_fall$YEAR
nyrs = length(years)
ages = c(0:15)
ages.lab = as.character(ages)
n_ages = length(ages)

#calculate proportion-at-age 
acomp.obs = GOM_agecomps_matrix_madmf_fall %>% select(4:19)
names(acomp.obs) <- as.numeric(ages)
catch.yrs = nrow(GOM_agecomps_matrix_madmf_fall)

#plot
my.title <- paste0("Catch-at-Age for GOM MADMF Fall Survey")
  origpar <- par(no.readonly = TRUE)
  par(mar=c(4,4,2,2), oma=c(1,1,1,1), mfrow=c(1,1))
  scale.catch.obs <- 0.1
  z3 <- as.matrix(acomp.obs) * scale.catch.obs

  plot(ages, rev(ages),  xlim = range(ages), ylim = c(years[nyrs],(years[1]-2)), xlab = "Age", ylab = "", type = "n", axes=FALSE)
  axis(1, at= ages, lab = ages.lab)
  axis(2, at = rev(years), lab = rev(years), cex.axis=0.75, las=1)
  box()
  abline(h=years, col="lightgray")
  segments(x0=ages, y0=rep(years[1],n_ages), x1=ages, y1=rep(years[nyrs],n_ages), col = "lightgray", lty = 1)
  if (length(catch.yrs)>0) for (j in 1:nyrs) points(ages, rep(years[j], n_ages), cex=z3[j,], col="black", #bg = bubble.col,
                                                    pch = 21)

  bubble.legend1 <- c(0.05,0.2,0.4)
  bubble.legend2 <- bubble.legend1 * scale.catch.obs
  legend("topright", xpd=TRUE, legend=bubble.legend1, pch=rep(21, 3), pt.cex=bubble.legend2, horiz=T , col='black', pt.bg = bubble.col)
  title (my.title, outer=TRUE, line=-1)
  par(origpar)
```

#Index catch-at-age code for SNEMA
```{r}
bubble.col = "white"

SNEMA_agecomps <- read.csv("~/Desktop/StockAssessmentPracticum/WF_NaturalMortality/STOCKEFF_SV_172905_SNEMA_NONE_strat_mean_age.csv")

#different surveys -> NMFS BTS + MADMF

#NMFS BTS

#turn file into asap matrix format 
SNEMA_agecomps_filter <- na.omit(SNEMA_agecomps)

SNEMA_agecomps_matrix_nmfs <- SNEMA_agecomps_filter %>%
  filter(grepl("NMFS", SURVEY)) %>%
select(STOCK_ABBREV, SEASON, YEAR, AGE, NO_AT_AGE) %>%
  pivot_wider(
    names_from = AGE,
    values_from = NO_AT_AGE
  )
col_nums <- sort(as.numeric(names(SNEMA_agecomps_matrix_nmfs[4:19])))
SNEMA_agecomps_matrix_nmfs <- SNEMA_agecomps_matrix_nmfs %>%
  select(STOCK_ABBREV, SEASON, YEAR, all_of(as.character(col_nums)))

#PLOT AGE COMPS BY SEASON 

#SNEMA spring 

#arrange data 
SNEMA_agecomps_matrix_nmfs_spring <- SNEMA_agecomps_matrix_nmfs %>% filter(SEASON == "SPRING") %>%
  mutate(across(everything(), ~replace(., is.na(.), 0)))
years = SNEMA_agecomps_matrix_nmfs_spring$YEAR
nyrs = length(years)
#ages = c(0:13,15,18)
ages = c(0:7)
ages.lab = c(as.character(ages[1:7]), "7+")
n_ages = length(ages)

#calculate proportion-at-age 
#acomp.obs = SNEMA_agecomps_matrix_nmfs_spring %>% select(4:19)

acomp.obs = SNEMA_agecomps_matrix_nmfs_spring %>% mutate('7' = rowSums(SNEMA_agecomps_matrix_nmfs_spring[, 11:19], na.rm = TRUE)) %>% select(4:11)
names(acomp.obs) <- as.numeric(ages)
catch.yrs = nrow(SNEMA_agecomps_matrix_nmfs_spring)

#plot
my.title <- paste0("NEFSC Bottom Trawl Survey SNEMA Spring Abundance-at-Age")
  origpar <- par(no.readonly = TRUE)
  par(mar=c(4,4,2,2), oma=c(1,1,1,1), mfrow=c(1,1))
  scale.catch.obs <- 2
  z3 <- as.matrix(acomp.obs) * scale.catch.obs

  plot(ages, rev(ages),  xlim = range(ages), ylim = c(years[nyrs],(years[1]-2)), xlab = "Age", ylab = "", type = "n", axes=FALSE)
  axis(1, at= ages, lab = ages.lab)
  axis(2, at = rev(years), lab = rev(years), cex.axis=0.75, las=1)
  box()
  abline(h=years, col="lightgray")
  segments(x0=ages, y0=rep(years[1],n_ages), x1=ages, y1=rep(years[nyrs],n_ages), col = "lightgray", lty = 1)
  if (length(catch.yrs)>0) for (j in 1:nyrs) points(ages, rep(years[j], n_ages), cex=z3[j,], col="black", #bg = bubble.col,
                                                    pch = 21)

  bubble.legend1 <- c(0.05,0.2,0.4)
  bubble.legend2 <- bubble.legend1 * scale.catch.obs
  legend("topright", xpd=TRUE, legend=bubble.legend1, pch=rep(21, 3), pt.cex=bubble.legend2, horiz=T , col='black', pt.bg = bubble.col)
  title (my.title, outer=TRUE, line=-1)
  par(origpar)
  
#SNEMA Fall
  
#arrange data 
SNEMA_agecomps_matrix_nmfs_fall <- SNEMA_agecomps_matrix_nmfs %>% filter(SEASON == "FALL") %>%
  mutate(across(everything(), ~replace(., is.na(.), 0)))
years = SNEMA_agecomps_matrix_nmfs_fall$YEAR
nyrs = length(years)
#ages = c(0:13,15, 18)
ages = c(0:7)
ages.lab = c(as.character(ages[1:7]), "7+")
n_ages = length(ages)

#calculate proportion-at-age 
acomp.obs = SNEMA_agecomps_matrix_nmfs_fall %>% select(4:19)

acomp.obs = SNEMA_agecomps_matrix_nmfs_fall %>% mutate('7' = rowSums(SNEMA_agecomps_matrix_nmfs_fall[, 11:19], na.rm = TRUE)) %>% select(4:11)
names(acomp.obs) <- as.numeric(ages)
catch.yrs = nrow(SNEMA_agecomps_matrix_nmfs_fall)

#plot
my.title <- paste0("NEFSC Bottom Trawl Survey SNEMA Fall Abundance-at-Age")
  origpar <- par(no.readonly = TRUE)
  par(mar=c(4,4,2,2), oma=c(1,1,1,1), mfrow=c(1,1))
  scale.catch.obs <- 2
  z3 <- as.matrix(acomp.obs) * scale.catch.obs

  plot(ages, rev(ages),  xlim = range(ages), ylim = c(years[nyrs],(years[1]-2)), xlab = "Age", ylab = "", type = "n", axes=FALSE)
  axis(1, at= ages, lab = ages.lab)
  axis(2, at = rev(years), lab = rev(years), cex.axis=0.75, las=1)
  box()
  abline(h=years, col="lightgray")
  segments(x0=ages, y0=rep(years[1],n_ages), x1=ages, y1=rep(years[nyrs],n_ages), col = "lightgray", lty = 1)
  if (length(catch.yrs)>0) for (j in 1:nyrs) points(ages, rep(years[j], n_ages), cex=z3[j,], col="black", #bg = bubble.col,
                                                    pch = 21)

  bubble.legend1 <- c(0.05,0.2,0.4)
  bubble.legend2 <- bubble.legend1 * scale.catch.obs
  legend("topright", xpd=TRUE, legend=bubble.legend1, pch=rep(21, 3), pt.cex=bubble.legend2, horiz=T , col='black', pt.bg = bubble.col)
  title (my.title, outer=TRUE, line=-1)
  par(origpar)
  
#MADMF
SNEMA_agecomps_filter <- na.omit(SNEMA_agecomps)

SNEMA_agecomps_matrix_madmf <- SNEMA_agecomps_filter %>%
  filter(grepl("MADMF", SURVEY)) %>%
select(STOCK_ABBREV, SEASON, YEAR, AGE, NO_AT_AGE) %>%
  pivot_wider(
    names_from = AGE,
    values_from = NO_AT_AGE
  )
col_nums <- sort(as.numeric(names(SNEMA_agecomps_matrix_madmf[4:18])))
SNEMA_agecomps_matrix_madmf <- SNEMA_agecomps_matrix_madmf %>%
  select(STOCK_ABBREV, SEASON, YEAR, all_of(as.character(col_nums)))

#PLOT AGE COMPS BY SEASON 

#SNEMA spring 

#arrange data 
SNEMA_agecomps_matrix_madmf_spring <- SNEMA_agecomps_matrix_madmf %>% filter(SEASON == "SPRING") %>%
  mutate(across(everything(), ~replace(., is.na(.), 0)))
years = SNEMA_agecomps_matrix_madmf_spring$YEAR
nyrs = length(years)
ages = c(1:14,16)
ages.lab = as.character(ages)
n_ages = length(ages)

#calculate proportion-at-age 
acomp.obs = SNEMA_agecomps_matrix_madmf_spring %>% select(4:18)
names(acomp.obs) <- as.numeric(ages)
catch.yrs = nrow(SNEMA_agecomps_matrix_madmf_spring)

#plot
my.title <- paste0("Catch-At-Age for SNEMA MADMF Spring Survey")
  origpar <- par(no.readonly = TRUE)
  par(mar=c(4,4,2,2), oma=c(1,1,1,1), mfrow=c(1,1))
  scale.catch.obs <- 0.5
  z3 <- as.matrix(acomp.obs) * scale.catch.obs

  plot(ages, rev(ages),  xlim = range(ages), ylim = c(years[nyrs],(years[1]-2)), xlab = "Age", ylab = "", type = "n", axes=FALSE)
  axis(1, at= ages, lab = ages.lab)
  axis(2, at = rev(years), lab = rev(years), cex.axis=0.75, las=1)
  box()
  abline(h=years, col="lightgray")
  segments(x0=ages, y0=rep(years[1],n_ages), x1=ages, y1=rep(years[nyrs],n_ages), col = "lightgray", lty = 1)
  if (length(catch.yrs)>0) for (j in 1:nyrs) points(ages, rep(years[j], n_ages), cex=z3[j,], col="black", #bg = bubble.col,
                                                    pch = 21)

  bubble.legend1 <- c(0.05,0.2,0.4)
  bubble.legend2 <- bubble.legend1 * scale.catch.obs
  legend("topright", xpd=TRUE, legend=bubble.legend1, pch=rep(21, 3), pt.cex=bubble.legend2, horiz=T , col='black', pt.bg = bubble.col)
  title (my.title, outer=TRUE, line=-1)
  par(origpar)
```

#CATCH CURVES

#GB catch curve
```{r}
#SPRING NMFS 
Dynamic_Z_GB_NMFS_Spring <- list()

for (i in c(1980:2023)){
GB_agecomps_nmfs_spring <- GB_agecomps_filter %>% 
  filter(grepl("NMFS", SURVEY)) %>%
select(STOCK_ABBREV, SEASON, YEAR, AGE, NO_AT_AGE) %>% 
  filter(SEASON == "SPRING")

GB_agecomps_testyear1 <- GB_agecomps_nmfs_spring %>% filter(YEAR == i, AGE == 3) 
GB_agecomps_testyear2 <- GB_agecomps_nmfs_spring %>% filter(YEAR == i+1, AGE == 4)
if(i == 2017) next
if(i == 2019) next
if(i == 2020) next 
ChangeinNo <- GB_agecomps_testyear2$NO_AT_AGE-GB_agecomps_testyear1$NO_AT_AGE

GB_agecomps_test_both <- rbind(GB_agecomps_testyear1, GB_agecomps_testyear2)
cc <- lm(log(NO_AT_AGE)~AGE, data = GB_agecomps_test_both)
ChangeinZ = coef(cc)[2]

Dynamic_Z_GB_NMFS_Spring[[i]] <- data.frame(year = as.double(paste(GB_agecomps_testyear1$YEAR)), ChangeZ = ChangeinZ, ChangeZAbs = ChangeinZ*-1, ChangeinNoAtAge = ChangeinNo, ChangeinNoAtAgeAbs = ChangeinNo*-1, Stock = paste(GB_agecomps_nmfs_spring[1,]$STOCK_ABBREV), Season = paste(GB_agecomps_nmfs_spring[1,]$SEASON)) 
}

Dynamic_Z_GB_NMFS_Spring_compile <- rbindlist(Dynamic_Z_GB_NMFS_Spring)
Dynamic_Z_GB_NMFS_Spring_compile

#FALL NMFS
Dynamic_Z_GB_NMFS_Fall <- list()

for (i in c(1980:2023)){
GB_agecomps_nmfs_Fall <- GB_agecomps_filter %>% 
  filter(grepl("NMFS", SURVEY)) %>%
select(STOCK_ABBREV, SEASON, YEAR, AGE, NO_AT_AGE) %>% 
  filter(SEASON == "FALL")

GB_agecomps_testyear1 <- GB_agecomps_nmfs_Fall %>% filter(YEAR == i, AGE == 3) 
GB_agecomps_testyear2 <- GB_agecomps_nmfs_Fall %>% filter(YEAR == i+1, AGE == 4)
if(i == 1989) next
if(i == 2019) next
if(i == 2020) next
ChangeinNo <- GB_agecomps_testyear2$NO_AT_AGE-GB_agecomps_testyear1$NO_AT_AGE

GB_agecomps_test_both <- rbind(GB_agecomps_testyear1, GB_agecomps_testyear2)
cc <- lm(log(NO_AT_AGE)~AGE, data = GB_agecomps_test_both)
ChangeinZ = coef(cc)[2]

Dynamic_Z_GB_NMFS_Fall[[i]] <- data.frame(year = as.double(paste(GB_agecomps_testyear1$YEAR)), ChangeZ = ChangeinZ, ChangeZAbs = ChangeinZ*-1, ChangeinNoAtAge = ChangeinNo, ChangeinNoAtAgeAbs = ChangeinNo*-1, Stock = paste(GB_agecomps_nmfs_Fall[1,]$STOCK_ABBREV), Season = paste(GB_agecomps_nmfs_Fall[1,]$SEASON)) 
}

Dynamic_Z_GB_NMFS_Fall_compile <- rbindlist(Dynamic_Z_GB_NMFS_Fall)
Dynamic_Z_GB_NMFS_Fall_compile <- na.omit(Dynamic_Z_GB_NMFS_Fall_compile)
Dynamic_Z_GB_NMFS_Fall_compile

#mean(Dynamic_Z_GB_NMFS_Spring_compile$ChangeZAbs)
#median(Dynamic_Z_GB_NMFS_Spring_compile$ChangeZAbs)

#mean(Dynamic_Z_GB_NMFS_Fall_compile$ChangeZAbs)
#median(Dynamic_Z_GB_NMFS_Fall_compile$ChangeZAbs)

#number of negatives spring: 18

#number of negatives fall: 16

#way too many negatives 
```

#GOM catch curve 
```{r}
#SPRING NMFS 
Dynamic_Z_GOM_NMFS_Spring <- list()

for (i in c(1979:2023)){
GOM_agecomps_nmfs_spring <- GOM_agecomps_filter %>% 
  filter(grepl("NMFS", SURVEY)) %>%
select(STOCK_ABBREV, SEASON, YEAR, AGE, NO_AT_AGE) %>% 
  filter(SEASON == "SPRING")

GOM_agecomps_testyear1 <- GOM_agecomps_nmfs_spring %>% filter(YEAR == i, AGE == 3) 
GOM_agecomps_testyear2 <- GOM_agecomps_nmfs_spring %>% filter(YEAR == i+1, AGE == 4)
if(i == 2019) next
if(i == 2020) next 
if(i == 2022) next
if(i == 2023) next
ChangeinNo <- GOM_agecomps_testyear2$NO_AT_AGE-GOM_agecomps_testyear1$NO_AT_AGE

GOM_agecomps_test_both <- rbind(GOM_agecomps_testyear1, GOM_agecomps_testyear2)
cc <- lm(log(NO_AT_AGE)~AGE, data = GOM_agecomps_test_both)
ChangeinZ = coef(cc)[2]

Dynamic_Z_GOM_NMFS_Spring[[i]] <- data.frame(year = as.double(paste(GOM_agecomps_testyear1$YEAR)), ChangeZ = ChangeinZ, ChangeZAbs = ChangeinZ*-1, ChangeinNoAtAge = ChangeinNo, ChangeinNoAtAgeAbs = ChangeinNo*-1, Stock = paste(GOM_agecomps_nmfs_spring[1,]$STOCK_ABBREV), Season = paste(GOM_agecomps_nmfs_spring[1,]$SEASON)) 
}

Dynamic_Z_GOM_NMFS_Spring_compile <- rbindlist(Dynamic_Z_GOM_NMFS_Spring)
Dynamic_Z_GOM_NMFS_Spring_compile

#FALL NMFS
Dynamic_Z_GOM_NMFS_Fall <- list()

for (i in c(1979:2023)){
GOM_agecomps_nmfs_Fall <- GOM_agecomps_filter %>% 
  filter(grepl("NMFS", SURVEY)) %>%
select(STOCK_ABBREV, SEASON, YEAR, AGE, NO_AT_AGE) %>% 
  filter(SEASON == "FALL")

GOM_agecomps_testyear1 <- GOM_agecomps_nmfs_Fall %>% filter(YEAR == i, AGE == 3) 
GOM_agecomps_testyear2 <- GOM_agecomps_nmfs_Fall %>% filter(YEAR == i+1, AGE == 4)
if(i == 2019) next
if(i == 2020) next
ChangeinNo <- GOM_agecomps_testyear2$NO_AT_AGE-GOM_agecomps_testyear1$NO_AT_AGE

GOM_agecomps_test_both <- rbind(GOM_agecomps_testyear1, GOM_agecomps_testyear2)
cc <- lm(log(NO_AT_AGE)~AGE, data = GOM_agecomps_test_both)
ChangeinZ = coef(cc)[2]

Dynamic_Z_GOM_NMFS_Fall[[i]] <- data.frame(year = as.double(paste(GOM_agecomps_testyear1$YEAR)), ChangeZ = ChangeinZ, ChangeZAbs = ChangeinZ*-1, ChangeinNoAtAge = ChangeinNo, ChangeinNoAtAgeAbs = ChangeinNo*-1, Stock = paste(GOM_agecomps_nmfs_Fall[1,]$STOCK_ABBREV), Season = paste(GOM_agecomps_nmfs_Fall[1,]$SEASON)) 
}

Dynamic_Z_GOM_NMFS_Fall_compile <- rbindlist(Dynamic_Z_GOM_NMFS_Fall)
Dynamic_Z_GOM_NMFS_Fall_compile <- na.omit(Dynamic_Z_GOM_NMFS_Fall_compile)
Dynamic_Z_GOM_NMFS_Fall_compile

#mean(Dynamic_Z_GOM_NMFS_Spring_compile$ChangeZAbs)
#median(Dynamic_Z_GOM_NMFS_Spring_compile$ChangeZAbs)

#mean(Dynamic_Z_GOM_NMFS_Fall_compile$ChangeZAbs)
#median(Dynamic_Z_GOM_NMFS_Fall_compile$ChangeZAbs)

#number of negatives spring: 10

#number of negatives fall: 7
```

#SNEMA catch curve 
```{r}
Dynamic_Z_SNEMA_NMFS_Spring <- list()

for (i in c(1980:2023)){
SNEMA_agecomps_nmfs_spring <- SNEMA_agecomps_filter %>% 
  filter(grepl("NMFS", SURVEY)) %>%
select(STOCK_ABBREV, SEASON, YEAR, AGE, NO_AT_AGE) %>% 
  filter(SEASON == "SPRING")

SNEMA_agecomps_testyear1 <- SNEMA_agecomps_nmfs_spring %>% filter(YEAR == i, AGE == 3) 
SNEMA_agecomps_testyear2 <- SNEMA_agecomps_nmfs_spring %>% filter(YEAR == i+1, AGE == 4)
ChangeinNo <- SNEMA_agecomps_testyear2$NO_AT_AGE-SNEMA_agecomps_testyear1$NO_AT_AGE

SNEMA_agecomps_test_both <- rbind(SNEMA_agecomps_testyear1, SNEMA_agecomps_testyear2)
cc <- lm(log(NO_AT_AGE)~AGE, data = SNEMA_agecomps_test_both)
ChangeinZ = coef(cc)[2]

Dynamic_Z_SNEMA_NMFS_Spring[[i]] <- data.frame(year = as.double(paste(SNEMA_agecomps_testyear1$YEAR)), ChangeZ = ChangeinZ, ChangeZAbs = ChangeinZ*-1, ChangeinNoAtAge = ChangeinNo, ChangeinNoAtAgeAbs = ChangeinNo*-1, Stock = paste(SNEMA_agecomps_nmfs_spring[1,]$STOCK_ABBREV), Season = paste(SNEMA_agecomps_nmfs_spring[1,]$SEASON)) 
}

Dynamic_Z_SNEMA_NMFS_Spring_compile <- rbindlist(Dynamic_Z_SNEMA_NMFS_Spring)
Dynamic_Z_SNEMA_NMFS_Spring_compile

#FALL NMFS
Dynamic_Z_SNEMA_NMFS_Fall <- list()

for (i in c(1980:2023)){
SNEMA_agecomps_nmfs_Fall <- SNEMA_agecomps_filter %>% 
  filter(grepl("NMFS", SURVEY)) %>%
select(STOCK_ABBREV, SEASON, YEAR, AGE, NO_AT_AGE) %>% 
  filter(SEASON == "FALL")

SNEMA_agecomps_testyear1 <- SNEMA_agecomps_nmfs_Fall %>% filter(YEAR == i, AGE == 3) 
SNEMA_agecomps_testyear2 <- SNEMA_agecomps_nmfs_Fall %>% filter(YEAR == i+1, AGE == 4)
if(i == 2019) next
if(i == 2020) next
ChangeinNo <- SNEMA_agecomps_testyear2$NO_AT_AGE-SNEMA_agecomps_testyear1$NO_AT_AGE

SNEMA_agecomps_test_both <- rbind(SNEMA_agecomps_testyear1, SNEMA_agecomps_testyear2)
cc <- lm(log(NO_AT_AGE)~AGE, data = SNEMA_agecomps_test_both)
ChangeinZ = coef(cc)[2]

Dynamic_Z_SNEMA_NMFS_Fall[[i]] <- data.frame(year = as.double(paste(SNEMA_agecomps_testyear1$YEAR)), ChangeZ = ChangeinZ, ChangeZAbs = ChangeinZ*-1, ChangeinNoAtAge = ChangeinNo, ChangeinNoAtAgeAbs = ChangeinNo*-1, Stock = paste(SNEMA_agecomps_nmfs_Fall[1,]$STOCK_ABBREV), Season = paste(SNEMA_agecomps_nmfs_Fall[1,]$SEASON)) 
}

Dynamic_Z_SNEMA_NMFS_Fall_compile <- rbindlist(Dynamic_Z_SNEMA_NMFS_Fall)
Dynamic_Z_SNEMA_NMFS_Fall_compile <- na.omit(Dynamic_Z_SNEMA_NMFS_Fall_compile)
Dynamic_Z_SNEMA_NMFS_Fall_compile

#View(Dynamic_Z_SNEMA_NMFS_Spring_compile)

#mean(Dynamic_Z_SNEMA_NMFS_Spring_compile$ChangeZAbs)
#median(Dynamic_Z_SNEMA_NMFS_Spring_compile$ChangeZAbs)

#mean(Dynamic_Z_SNEMA_NMFS_Fall_compile$ChangeZAbs)
#median(Dynamic_Z_SNEMA_NMFS_Fall_compile$ChangeZAbs)

#number of negatives spring: 3

#number of negatives fall: 4
```

#READ IN WINTER FLOUNDER, PREDATORS AND COMPETITORS FOR CATCH RATE
```{r}
NEFSC_species_shipall_catchrate_ <- readRDS("~/Desktop/StockAssessmentPracticum/WF_NaturalMortality/NEFSC_species_allship_catchrate_.rds")

GB_springnumber <- NEFSC_species_shipall_catchrate_$GB_spring %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("num$|year$|Ship$"
))
GB_springnumber

GB_springbiomass <- NEFSC_species_shipall_catchrate_$GB_spring %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("wt$|year$|Ship$"
))
GB_springbiomass

GB_fallnumber <- NEFSC_species_shipall_catchrate_$GB_fall %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("num$|year$|Ship$"
))
GB_fallnumber

GB_fallbiomass <- NEFSC_species_shipall_catchrate_$GB_fall %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("wt$|year$|Ship$"
))
GB_fallbiomass

GOM_springnumber <- NEFSC_species_shipall_catchrate_$GOM_spring %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("num$|year$|Ship$"
))
GOM_springnumber

GOM_springbiomass <- NEFSC_species_shipall_catchrate_$GOM_spring %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("wt$|year$|Ship$"
))
GOM_springbiomass

GOM_fallnumber <- NEFSC_species_shipall_catchrate_$GOM_fall %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("num$|year$|Ship$"
))
GOM_fallnumber

GOM_fallbiomass <- NEFSC_species_shipall_catchrate_$GOM_fall %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("wt$|year$|Ship$"
))
GOM_fallbiomass

SNEMA_springnumber <- NEFSC_species_shipall_catchrate_$SNEMA_spring %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("num$|year$|Ship$"
)) %>% filter(!year %in% c(2023))
SNEMA_springnumber

SNEMA_springbiomass <- NEFSC_species_shipall_catchrate_$SNEMA_spring %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("wt$|year$|Ship$"
)) %>% filter(!year %in% c(2023))
SNEMA_springbiomass

SNEMA_fallnumber <- NEFSC_species_shipall_catchrate_$SNEMA_fall %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("num$|year$|Ship$"
)) %>% filter(!year %in% c(2017))
SNEMA_fallnumber

SNEMA_fallbiomass <- NEFSC_species_shipall_catchrate_$SNEMA_fall %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("wt$|year$|Ship$"
)) %>% filter(!year %in% c(2017))
SNEMA_fallbiomass
```

#READ IN PREDATORS AND COMPETITORS FOR MORTALITY RATE
```{r}
NEFSC_species_shipall_catchrate_ <- readRDS("~/Desktop/StockAssessmentPracticum/WF_NaturalMortality/NEFSC_species_allship_catchrate_.rds")

GOM_springnumber_mortality <- NEFSC_species_shipall_catchrate_$GOM_spring %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("num$|year$|Ship$"
))
GOM_springnumber_species_compiler <- GOM_springnumber_mortality %>% inner_join(Dynamic_Z_GOM_NMFS_Spring_compile, by = "year")
GOM_springnumber_species_compiler

GOM_springbiomass_mortality <- NEFSC_species_shipall_catchrate_$GOM_spring %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("wt$|year$|Ship$"
))
GOM_springbiomass_species_compiler <- GOM_springbiomass_mortality %>% inner_join(Dynamic_Z_GOM_NMFS_Spring_compile, by = "year")
GOM_springbiomass_species_compiler

GOM_fallnumber_mortality <- NEFSC_species_shipall_catchrate_$GOM_fall %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("num$|year$|Ship$"
))
GOM_fallnumber_species_compiler <- GOM_fallnumber_mortality %>% inner_join(Dynamic_Z_GOM_NMFS_Fall_compile, by = "year")
GOM_fallnumber_species_compiler

GOM_fallbiomass_mortality <- NEFSC_species_shipall_catchrate_$GOM_fall %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("wt$|year$|Ship$"
))
GOM_fallbiomass_species_compiler <- GOM_fallbiomass_mortality %>% inner_join(Dynamic_Z_GOM_NMFS_Fall_compile, by = "year")
GOM_fallbiomass_species_compiler

SNEMA_springnumber_mortality <- NEFSC_species_shipall_catchrate_$SNEMA_spring %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("num$|year$|Ship$"
))
SNEMA_springnumber_species_compiler <- SNEMA_springnumber_mortality %>% inner_join(Dynamic_Z_SNEMA_NMFS_Spring_compile, by = "year") %>% 
  #filter(year != 2023)
  filter(!year %in% c(2022, 2023))
SNEMA_springnumber_species_compiler

SNEMA_springbiomass_mortality <- NEFSC_species_shipall_catchrate_$SNEMA_spring %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("wt$|year$|Ship$"
))
SNEMA_springbiomass_species_compiler <- SNEMA_springbiomass_mortality %>% inner_join(Dynamic_Z_SNEMA_NMFS_Spring_compile, by = "year") %>% 
  #filter(year != 2023)
  filter(!year %in% c(2022, 2023))
SNEMA_springbiomass_species_compiler

SNEMA_fallnumber_mortality <- NEFSC_species_shipall_catchrate_$SNEMA_fall %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("num$|year$|Ship$"
))
SNEMA_fallnumber_species_compiler <- SNEMA_fallnumber_mortality %>% inner_join(Dynamic_Z_SNEMA_NMFS_Fall_compile, by = "year") %>% 
  #filter(year != 2017)
  filter(!year %in% c(2016, 2017))
SNEMA_fallnumber_species_compiler

SNEMA_fallbiomass_mortality <- NEFSC_species_shipall_catchrate_$SNEMA_fall %>% mutate(Ship = as.factor(ifelse(year %in% c(1963:2008), "ALB", "BIG"))) %>% select(matches("wt$|year$|Ship$"
))
SNEMA_fallbiomass_species_compiler <- SNEMA_fallbiomass_mortality %>% inner_join(Dynamic_Z_SNEMA_NMFS_Fall_compile, by = "year") %>% 
  #filter(year != 2017)
  filter(!year %in% c(2016, 2017))
SNEMA_fallbiomass_species_compiler

```

#Power analysis
```{r}
library(pwr)

k <- 1
alpha <- 0.05
power <- 0.8 

R2_values <- seq(0.02, 0.5, by = 0.01)  

sample_sizes <- sapply(R2_values, function(R2) {
  f2 <- R2 / (1 - R2)
  result <- pwr.f2.test(u = k, f2 = f2, sig.level = alpha, power = power)
  return(ceiling(result$v + k + 1))  
})
sample_size_df <- data.frame(R2 = R2_values, SampleSize = sample_sizes)
print(sample_size_df)

#catch rate
nrow(GB_springnumber) #56
nrow(GB_fallnumber) #61

nrow(GOM_springnumber)  #55
nrow(GOM_fallnumber) #61

nrow(SNEMA_springnumber) #56
nrow(SNEMA_fallnumber) #60

#mortality rate
nrow(GOM_springnumber_species_compiler) #41 
nrow(GOM_fallnumber_species_compiler) #43 

nrow(SNEMA_springnumber_species_compiler) #42 
nrow(SNEMA_fallnumber_species_compiler) #40 
```

#Do fixed group-level intercept regression w/rapid correlation function - catch rate
```{r}
#full function with plot

group_reg_allship <- function(data, target_var, other_vars, group_var) {
  library(ggplot2)
  library(dplyr)
  
  cor_results <- list()
  
  for (var in other_vars) {
    formula <- as.formula(paste(target_var, "~", var, "+", group_var)) #slope or intercept
    test_result <- lm(formula, data = data)
    mod_sum <- summary(test_result)
    f_stat <- mod_sum$fstatistic
    
    if (is.null(f_stat)) { next }
    
    #p value of species
    #p_value <- mod_sum$coefficients[11] #this way is honestly fine for the model in its current form unless you add more predictors
    coef_names <- rownames(mod_sum$coefficients)
    var_coef_name <- coef_names[startsWith(coef_names, paste0(var))]
    if (length(var_coef_name) == 0) { 
      p_value <- NA
      coef_val <- 0 
      } 
    else { 
    p_value <- mod_sum$coefficients[var_coef_name, "Pr(>|t|)"] 
    coef_val <- coef(test_result)[var_coef_name] 
    }
    
    #p value of ship 
    #p_value_ship <- mod_sum$coefficients[12]
    coef_names <- rownames(mod_sum$coefficients)
    group_var_name <- coef_names[startsWith(coef_names, paste0(group_var))] 
    p_value_ship <- mod_sum$coefficients[group_var_name, "Pr(>|t|)"]
    
    #model p value
    p_value_mod <- pf(f_stat[1], f_stat[2], f_stat[3], lower.tail = FALSE)
    
    #coefficient
    coef_val <- coef(test_result)[var]
    
    #R2 
    r_squared = mod_sum$r.squared
    
    #get rid of error in cases where p value is low but coef_val = NA because all values for a species are 0 
    coef_val[is.na(coef_val)] <- 0
    
    cor_results[[length(cor_results) + 1]] <- data.frame(
      Target_Var = target_var,
      Other_Var = var,
      R_Squared = mod_sum$r.squared,
      P_Value = p_value,
      P_Value_ship = p_value_ship,
      P_Value_mod = p_value_mod,
      Coefficient = coef_val
    )
    
    # Plot only if p-value is significant
    if (!is.na(p_value) && p_value <= 0.05 && coef_val < 0) {
      
      # if(!is.na(p_value)) {
      
      # Add predicted values to data
      data_with_pred <- data %>%
        mutate(predicted = predict(test_result))
      
      #clean variable names for plotting 
      clean_var <- tools::toTitleCase( gsub("_", " ", sub("(_num|_wt)$", "", var)) )
      if (grepl("Goosefish", clean_var, ignore.case = TRUE)) {clean_var <- "Monkfish"}
    
      data_name <- deparse(substitute(data))
      
  if (grepl("number", data_name, ignore.case = TRUE)) { datatype <- "Number/tow" } else 
  if (grepl("biomass", data_name, ignore.case = TRUE)) { datatype <- "kg/tow" }
      
  if (grepl("spring", data_name, ignore.case = TRUE)) { season <- "Spring" } else 
  if (grepl("fall", data_name, ignore.case = TRUE)) { season <- "Fall" }
      
  if (grepl("GB_", data_name, ignore.case = TRUE)) { stockregion <- "GB" } else 
  if (grepl("GOM_", data_name, ignore.case = TRUE)) { stockregion <- "GOM" } else 
  if (grepl("SNEMA_", data_name, ignore.case = TRUE)) { stockregion <- "SNEMA" }

      #remove spaces between axis labels
      pred_label <- stringr::str_squish( paste(stockregion, season, clean_var, datatype) )
      
      #give name to winter flounder catch rate
  if (grepl("winter_flounder_num", target_var, ignore.case = TRUE)) { winterflounder_name <- "Winter Flounder Number/tow" } else 
  if (grepl("winter_flounder_wt", target_var, ignore.case = TRUE)) { winterflounder_name <- "Winter Flounder kg/tow" }
      
      #extract levels from group_var
      group_levels <- levels(factor(data[[group_var]]))

      # Create lm plot
      ggplot(data_with_pred, aes_string(x = var, y = target_var, color = group_var)) +
        geom_point() +
      #geom_text(aes(x = .data[[var]], y = .data[[target_var]], label = year),
      #           vjust = -0.7, size = 3, color = "black", inherit.aes = FALSE ) + 
        geom_line(aes(y = predicted), size = 1) +
        labs(
          subtitle = paste(" R² =", round(r_squared,3), "| p =", round(p_value, 3)),
          x = pred_label,
          y = winterflounder_name
        ) +
        theme_minimal() + 
  scale_color_manual( name = "Survey Ship", 
                      labels = c( "ALB" = "NOAAS Albatross", 
                                  "BIG" = "NOAAS Henry B. Bigelow"),
                      values = c( "ALB" = "#154360",
                                  "BIG" = "#FFC300")
                      ) + 
         theme( legend.position = c(0.95, 0.05), 
                 legend.justification = c("right", "bottom"), 
                 legend.background = element_rect(fill = alpha("white", 0.7), 
                                                  color = NA) )-> plot_obj

      #survey trend plot
      
      #match predator/competitor variables and winter flounder variable to linetype 
      
      linetype_names <- c("solid", "dashed")
      names(linetype_names) <- c(pred_label,winterflounder_name)
      
      #set automatic scaling of axes
    scale_factor <- max(data[[var]], na.rm = TRUE) /
      max(data[[target_var]], na.rm = TRUE)

      #plot
      ggplot(data, aes(x = year)) +
        geom_line(aes(y = .data[[var]], 
                      linetype = pred_label)) + 
        geom_point(aes(y = .data[[var]])) +
        geom_line(aes(y = .data[[target_var]] * scale_factor, 
                linetype = winterflounder_name)) +
        geom_point(aes(y = .data[[target_var]] * scale_factor)) + 
        scale_y_continuous(name = pred_label, 
                      sec.axis = sec_axis( ~ . / scale_factor, 
                      name = winterflounder_name
                                           ) ) + 
  theme_classic() + 
  xlab("Year") + 
  scale_linetype_manual( name = "Time Series",
                         values = linetype_names) + 
        theme( legend.position = c(0.5, 0.98), 
                 legend.justification = c("center", "top"), 
                 legend.background = element_rect(fill = alpha("white", 0.7), 
                                                  color = NA) ) -> plot_obj2

      #print(plot_obj)
      #print(plot_obj2)

      combined_plot <- plot_obj | plot_obj2 
      print(combined_plot)
     
      #change path as appropriate  
      ggsave(filename = file.path("~/Desktop/StockAssessmentPracticum/Figures/CatchRate", paste0("combined_plot_", stockregion,"_",season,"_", var, ".png")), 
             plot = combined_plot,
             width = 12, height = 5, dpi = 300)
    }
  }

  all_cor_results <- bind_rows(cor_results)
  return(all_cor_results)
}

#catch rate
GB_springnumber <- as_tibble(GB_springnumber) 
GB_springnumber$Ship <- as.factor(GB_springnumber$Ship)
target_var <- "winter_flounder_num"
other_var <- colnames(GB_springnumber[3:32])
group_var <- "Ship"
GB_spring_cor_results_allship <- group_reg_allship(GB_springnumber, target_var, other_var, group_var)
GB_spring_cor <- GB_spring_cor_results_allship %>% mutate(Scen = "GB_spring_lowcorrelation_number")
correlation_filter_GBspringnumber <-GB_spring_cor %>% filter(P_Value <= 0.05 & Coefficient < 0)
print(correlation_filter_GBspringnumber)

GB_springbiomass <- as_tibble(GB_springbiomass) 
GB_springbiomass$Ship <- as.factor(GB_springbiomass$Ship)
target_var <- "winter_flounder_wt"
other_var <- colnames(GB_springbiomass[3:32])
group_var <- "Ship"
GB_spring_cor_results_allship <- group_reg_allship(GB_springbiomass, target_var, other_var, group_var)
GB_spring_cor <- GB_spring_cor_results_allship %>% mutate(Scen = "GOM_spring_lowcorrelation_biomass")
correlation_filter_GBspringbiomass <-GB_spring_cor %>% filter(P_Value <= 0.05 & Coefficient < 0)
print(correlation_filter_GBspringbiomass)

GB_fallnumber <- as_tibble(GB_fallnumber) 
GB_fallnumber$Ship <- as.factor(GB_fallnumber$Ship)
target_var <- "winter_flounder_num"
other_var <- colnames(GB_fallnumber[3:32])
group_var <- "Ship"
GB_fall_cor_results_allship <- group_reg_allship(GB_fallnumber, target_var, other_var, group_var)
GB_fall_cor <- GB_fall_cor_results_allship %>% mutate(Scen = "GB_fall_lowcorrelation_number")
correlation_filter_GBfallnumber <-GB_fall_cor %>% filter(P_Value <= 0.05 & Coefficient < 0)
print(correlation_filter_GBfallnumber)

GB_fallbiomass <- as_tibble(GB_fallbiomass) 
GB_fallbiomass$Ship <- as.factor(GB_fallbiomass$Ship)
target_var <- "winter_flounder_wt"
other_var <- colnames(GB_fallbiomass[3:32])
group_var <- "Ship"
GB_fall_cor_results_allship <- group_reg_allship(GB_fallbiomass, target_var, other_var, group_var)
GB_fall_cor <- GB_fall_cor_results_allship %>% mutate(Scen = "GB_fall_lowcorrelation_biomass")
correlation_filter_GBfallbiomass <-GB_fall_cor %>% filter(P_Value <= 0.05 & Coefficient < 0)
print(correlation_filter_GBfallbiomass)

GOM_springnumber <- as_tibble(GOM_springnumber) 
GOM_springnumber$Ship <- as.factor(GOM_springnumber$Ship)
target_var <- "winter_flounder_num"
other_var <- colnames(GOM_springnumber[3:32])
group_var <- "Ship"
GOM_spring_cor_results_allship <- group_reg_allship(GOM_springnumber, target_var, other_var, group_var)
GOM_spring_cor <- GOM_spring_cor_results_allship %>% mutate(Scen = "GOM_spring_lowcorrelation_number")
correlation_filter_GOMspringnumber <-GOM_spring_cor %>% filter(P_Value <= 0.05 & Coefficient < 0)
print(correlation_filter_GOMspringnumber)

GOM_springbiomass <- as_tibble(GOM_springbiomass) 
GOM_springbiomass$Ship <- as.factor(GOM_springbiomass$Ship)
target_var <- "winter_flounder_wt"
other_var <- colnames(GOM_springbiomass[3:32])
group_var <- "Ship"
GOM_spring_cor_results_allship <- group_reg_allship(GOM_springbiomass, target_var, other_var, group_var)
GOM_spring_cor <- GOM_spring_cor_results_allship %>% mutate(Scen = "GOM_spring_lowcorrelation_biomass")
correlation_filter_GOMspringbiomass <-GOM_spring_cor %>% filter(P_Value <= 0.05 & Coefficient < 0)
print(correlation_filter_GOMspringbiomass)

GOM_fallnumber <- as_tibble(GOM_fallnumber) 
GOM_fallnumber$Ship <- as.factor(GOM_fallnumber$Ship)
target_var <- "winter_flounder_num"
other_var <- colnames(GOM_fallnumber[3:32])
group_var <- "Ship"
GOM_fall_cor_results_allship <- group_reg_allship(GOM_fallnumber, target_var, other_var, group_var)
GOM_fall_cor <- GOM_fall_cor_results_allship %>% mutate(Scen = "GOM_fall_lowcorrelation_number")
correlation_filter_GOMfallnumber <-GOM_fall_cor %>% filter(P_Value <= 0.05 & Coefficient < 0)
print(correlation_filter_GOMfallnumber)

GOM_fallbiomass <- as_tibble(GOM_fallbiomass) 
GOM_fallbiomass$Ship <- as.factor(GOM_fallbiomass$Ship)
target_var <- "winter_flounder_wt"
other_var <- colnames(GOM_fallbiomass[3:32])
group_var <- "Ship"
GOM_fall_cor_results_allship <- group_reg_allship(GOM_fallbiomass, target_var, other_var, group_var)
GOM_fall_cor <- GOM_fall_cor_results_allship %>% mutate(Scen = "GOM_fall_lowcorrelation_biomass")
correlation_filter_GOMfallbiomass <-GOM_fall_cor %>% filter(P_Value <= 0.05 & Coefficient < 0)
print(correlation_filter_GOMfallbiomass)

SNEMA_springnumber <- as_tibble(SNEMA_springnumber)
SNEMA_springnumber$Ship <- as.factor(SNEMA_springnumber$Ship)
target_var <- "winter_flounder_num"
other_var <- colnames(SNEMA_springnumber[3:32])
group_var <- "Ship"
SNEMA_spring_cor_results_allship <- group_reg_allship(SNEMA_springnumber, target_var, other_var, group_var)
SNEMA_spring_cor <- SNEMA_spring_cor_results_allship %>% mutate(Scen = "SNEMA_spring_lowcorrelation_number")
correlation_filter_SNEMAspringnumber <-SNEMA_spring_cor %>% filter(P_Value <= 0.05 & Coefficient < 0)
print(correlation_filter_SNEMAspringnumber)

SNEMA_springbiomass <- as_tibble(SNEMA_springbiomass) 
SNEMA_springbiomass$Ship <- as.factor(SNEMA_springbiomass$Ship)
target_var <- "winter_flounder_wt"
other_var <- colnames(SNEMA_springbiomass[3:32])
group_var <- "Ship"
SNEMA_spring_cor_results_allship <- group_reg_allship(SNEMA_springbiomass, target_var, other_var, group_var)
SNEMA_spring_cor <- SNEMA_spring_cor_results_allship %>% mutate(Scen = "SNEMA_spring_lowcorrelation_biomass")
correlation_filter_SNEMAspringbiomass <-SNEMA_spring_cor %>% filter(P_Value <= 0.05 & Coefficient < 0)
print(correlation_filter_SNEMAspringbiomass)

SNEMA_fallnumber <- as_tibble(SNEMA_fallnumber)
SNEMA_fallnumber$Ship <- as.factor(SNEMA_fallnumber$Ship)
target_var <- "winter_flounder_num"
other_var <- colnames(SNEMA_fallnumber[3:32])
group_var <- "Ship"
SNEMA_fall_cor_results_allship <- group_reg_allship(SNEMA_fallnumber, target_var, other_var, group_var)
SNEMA_fall_cor <- SNEMA_fall_cor_results_allship %>% mutate(Scen = "SNEMA_fall_lowcorrelation_number")
correlation_filter_SNEMAfallnumber <-SNEMA_fall_cor %>% filter(P_Value <= 0.05 & Coefficient < 0)
print(correlation_filter_SNEMAfallnumber)

SNEMA_fallbiomass <- as_tibble(SNEMA_fallbiomass)
SNEMA_fallbiomass$Ship <- as.factor(SNEMA_fallbiomass$Ship)
target_var <- "winter_flounder_wt"
other_var <- colnames(SNEMA_fallbiomass[3:32])
group_var <- "Ship"
SNEMA_fall_cor_results_allship <- group_reg_allship(SNEMA_fallbiomass, target_var, other_var, group_var)
SNEMA_fall_cor <- SNEMA_fall_cor_results_allship %>% mutate(Scen = "SNEMA_fall_lowcorrelation_biomass")
correlation_filter_SNEMAfallbiomass <-SNEMA_fall_cor %>% filter(P_Value <= 0.05 & Coefficient < 0)
print(correlation_filter_SNEMAfallbiomass)

combinedd <- rbind(correlation_filter_GBspringnumber, correlation_filter_GBfallnumber, correlation_filter_GBspringbiomass, correlation_filter_GBfallbiomass, correlation_filter_GOMspringnumber, correlation_filter_GOMfallnumber, correlation_filter_GOMspringbiomass, correlation_filter_GOMfallbiomass, correlation_filter_SNEMAspringnumber, correlation_filter_SNEMAfallnumber, correlation_filter_SNEMAspringbiomass, correlation_filter_SNEMAfallbiomass)
combinedd
```

#Do fixed group-level intercept regression w/rapid correlation function - mortality rate
```{r}
#full function with plot
group_reg_allship <- function(data, target_var, other_vars, group_var) {
  library(ggplot2)
  library(dplyr)

  cor_results <- list()

  for (var in other_vars) {
    formula <- as.formula(paste(target_var, "~", var, "+", group_var)) #slope or intercept
    test_result <- lm(formula, data = data)
    mod_sum <- summary(test_result)
    f_stat <- mod_sum$fstatistic
    
   #p value of species
   #p_value <- mod_sum$coefficients[11] #this way is honestly fine for the model in its current form unless you add more predictors
    coef_names <- rownames(mod_sum$coefficients)
    var_coef_name <- coef_names[startsWith(coef_names, paste0(var))]
    if (length(var_coef_name) == 0) { 
      p_value <- NA
      coef_val <- 0 
      } 
    else { 
    p_value <- mod_sum$coefficients[var_coef_name, "Pr(>|t|)"] 
    coef_val <- coef(test_result)[var_coef_name] 
    }
    
    #p value of ship 
    #p_value_ship <- mod_sum$coefficients[12]
    coef_names <- rownames(mod_sum$coefficients)
    group_var_name <- coef_names[startsWith(coef_names, paste0(group_var))] 
    p_value_ship <- mod_sum$coefficients[group_var_name, "Pr(>|t|)"]
    
    #model p value
    p_value_mod <- pf(f_stat[1], f_stat[2], f_stat[3], lower.tail = FALSE)
    
    #coefficient
    coef_val <- coef(test_result)[var]
    
    #R2 
    r_squared = mod_sum$r.squared
    
    #get rid of error in cases where p value is low but coef_val = NA because all values for a species are 0 
    coef_val[is.na(coef_val)] <- 0

    cor_results[[length(cor_results) + 1]] <- data.frame(
      Target_Var = target_var,
      Other_Var = var,
      R_Squared = r_squared,
      P_Value = p_value,
      P_Value_ship = p_value_ship,
      P_Value_mod = p_value_mod,
      Coefficient = coef_val
    )

    # Plot only if p-value is significant
   if (!is.na(p_value) && p_value <= 0.05 && coef_val > 0) {
    
    # if(!is.na(p_value)) {
     
      # Add predicted values to data
      data_with_pred <- data %>%
        mutate(predicted = predict(test_result))
      
      #clean variable names for plotting 
      clean_var <- tools::toTitleCase( gsub("_", " ", sub("(_num|_wt)$", "", var)) )
      if (grepl("Goosefish", clean_var, ignore.case = TRUE)) {clean_var <- "Monkfish"}
    
      data_name <- deparse(substitute(data))
      
  if (grepl("number", data_name, ignore.case = TRUE)) { datatype <- "Number/tow" } else 
  if (grepl("biomass", data_name, ignore.case = TRUE)) { datatype <- "kg/tow" }
      
  if (grepl("spring", data_name, ignore.case = TRUE)) { season <- "Spring" } else 
  if (grepl("fall", data_name, ignore.case = TRUE)) { season <- "Fall" }
      
  if (grepl("GB_", data_name, ignore.case = TRUE)) { stockregion <- "GB" } else 
  if (grepl("GOM_", data_name, ignore.case = TRUE)) { stockregion <- "GOM" } else 
  if (grepl("SNEMA_", data_name, ignore.case = TRUE)) { stockregion <- "SNEMA" }

      #remove spaces between axis labels
      pred_label <- stringr::str_squish( paste(stockregion, season, clean_var, datatype) )
      
      #give name to winter flounder Z
      winterflounder_name = "Winter Flounder Mortality Rate (Z)"
      
      #extract levels from group_var
      group_levels <- levels(factor(data[[group_var]]))
      
      # Create lm plot
      ggplot(data_with_pred, aes_string(x = var, y = target_var, color = group_var)) +
        geom_point() +
      #geom_text(aes(x = .data[[var]], y = .data[[target_var]], label = year),
      #           vjust = -0.7, size = 3, color = "black", inherit.aes = FALSE ) + 
        geom_line(aes(y = predicted), size = 1) +
        labs(
          subtitle = paste(" R² =", round(r_squared,3), "| p =", round(p_value, 3)),
          x = pred_label,
          y = winterflounder_name
        ) +
        theme_minimal() + 
  scale_color_manual( name = "Survey Ship", 
                      labels = c( "ALB" = "NOAAS Albatross", 
                                  "BIG" = "NOAAS Henry B. Bigelow"),
                      values = c( "ALB" = "#154360",
                                  "BIG" = "#FFC300")
                      ) + 
         theme( legend.position = c(0.95, 0.05), 
                 legend.justification = c("right", "bottom"), 
                 legend.background = element_rect(fill = alpha("white", 0.7), 
                                                  color = NA) ) -> plot_obj
      
      #survey trend plot
      
      #match predator/competitor variables and winter flounder variable to linetype 
      
      linetype_names <- c("solid", "dashed")
      names(linetype_names) <- c(pred_label,winterflounder_name)
      
      #set automatic scaling of axes
    scale_factor <- max(data[[var]], na.rm = TRUE) /
      max(data[[target_var]], na.rm = TRUE)

      #plot
      ggplot(data, aes(x = year)) +
        geom_line(aes(y = .data[[var]], 
                      linetype = pred_label)) + 
        geom_point(aes(y = .data[[var]])) +
        geom_line(aes(y = .data[[target_var]] * scale_factor, 
                linetype = winterflounder_name)) +
        geom_point(aes(y = .data[[target_var]] * scale_factor)) + 
        scale_y_continuous(name = pred_label, 
                      sec.axis = sec_axis( ~ . / scale_factor, 
                      name = winterflounder_name
                                           ) ) + 
  theme_classic() + 
  xlab("Year") + 
  scale_linetype_manual( name = "Time Series",
                         values = linetype_names) + 
        theme( legend.position = c(0.5, 0.98), 
                 legend.justification = c("center", "top"), 
                 legend.background = element_rect(fill = alpha("white", 0.7), 
                                                  color = NA) ) -> plot_obj2

      #print(plot_obj)
      #print(plot_obj2)
      
      combined_plot <- plot_obj | plot_obj2 
      print(combined_plot)
      
     #change path as appropriate
      ggsave(filename = file.path("~/Desktop/StockAssessmentPracticum/Figures/MortalityRate", paste0("combined_plot_", stockregion,"_",season,"_", var, ".png")), 
             plot = combined_plot,
             width = 12, height = 5, dpi = 300)
    }
  }

  all_cor_results <- bind_rows(cor_results)
  return(all_cor_results)
}

#catch curve 
GOM_springnumber_species_compiler <- as_tibble(GOM_springnumber_species_compiler)
GOM_springnumber_species_compiler$Ship <- as.factor(GOM_springnumber_species_compiler$Ship)
target_var <- "ChangeZAbs"
other_var <- colnames(GOM_springnumber_species_compiler[3:32])
group_var <- "Ship"
GOM_spring_cor_results_allship <- group_reg_allship(GOM_springnumber_species_compiler, target_var, other_var, group_var)
GOM_spring_cor <- GOM_spring_cor_results_allship %>% mutate(Scen = "GOM_spring_lowcorrelation_number")
correlation_filter_GOMspringnumber <-GOM_spring_cor %>% filter(P_Value <= 0.05 & Coefficient > 0)
print(correlation_filter_GOMspringnumber)

GOM_springbiomass_species_compiler <- as_tibble(GOM_springbiomass_species_compiler)
GOM_springbiomass_species_compiler$Ship <- as.factor(GOM_springbiomass_species_compiler$Ship)
target_var <- "ChangeZAbs"
other_var <- colnames(GOM_springbiomass_species_compiler[3:32])
group_var <- "Ship"
GOM_spring_cor_results_allship <- group_reg_allship(GOM_springbiomass_species_compiler, target_var, other_var, group_var)
GOM_spring_cor <- GOM_spring_cor_results_allship %>% mutate(Scen = "GOM_spring_lowcorrelation_biomass")
correlation_filter_GOMspringbiomass <-GOM_spring_cor %>% filter(P_Value <= 0.05 & Coefficient > 0)
print(correlation_filter_GOMspringbiomass)

GOM_fallnumber_species_compiler <- as_tibble(GOM_fallnumber_species_compiler) 
GOM_fallnumber_species_compiler$Ship <- as.factor(GOM_fallnumber_species_compiler$Ship)
target_var <- "ChangeZAbs"
other_var <- colnames(GOM_fallnumber_species_compiler[3:32])
group_var <- "Ship"
GOM_fall_cor_results_allship <- group_reg_allship(GOM_fallnumber_species_compiler, target_var, other_var, group_var)
GOM_fall_cor <- GOM_fall_cor_results_allship %>% mutate(Scen = "GOM_fall_lowcorrelation_number")
correlation_filter_GOMfallnumber <-GOM_fall_cor %>% filter(P_Value <= 0.05 & Coefficient > 0)
print(correlation_filter_GOMfallnumber)

GOM_fallbiomass_species_compiler <- as_tibble(GOM_fallbiomass_species_compiler) 
GOM_fallbiomass_species_compiler$Ship <- as.factor(GOM_fallbiomass_species_compiler$Ship)
target_var <- "ChangeZAbs"
other_var <- colnames(GOM_fallbiomass_species_compiler[3:32])
group_var <- "Ship"
GOM_fall_cor_results_allship <- group_reg_allship(GOM_fallbiomass_species_compiler, target_var, other_var, group_var)
GOM_fall_cor <- GOM_fall_cor_results_allship %>% mutate(Scen = "GOM_fall_lowcorrelation_biomass")
correlation_filter_GOMfallbiomass <-GOM_fall_cor %>% filter(P_Value <= 0.05 & Coefficient > 0)
print(correlation_filter_GOMfallbiomass)

SNEMA_springnumber_species_compiler <- as_tibble(SNEMA_springnumber_species_compiler)
SNEMA_springnumber_species_compiler$Ship <- as.factor(SNEMA_springnumber_species_compiler$Ship)
target_var <- "ChangeZAbs"
other_var <- colnames(SNEMA_springnumber_species_compiler[3:32])
group_var <- "Ship"
SNEMA_spring_cor_results_allship <- group_reg_allship(SNEMA_springnumber_species_compiler, target_var, other_var, group_var)
SNEMA_spring_cor <- SNEMA_spring_cor_results_allship %>% mutate(Scen = "SNEMA_spring_lowcorrelation_number")
correlation_filter_SNEMAspringnumber <-SNEMA_spring_cor %>% filter(P_Value <= 0.05 & Coefficient > 0)
print(correlation_filter_SNEMAspringnumber)

SNEMA_springbiomass_species_compiler <- as_tibble(SNEMA_springbiomass_species_compiler) 
SNEMA_springbiomass_species_compiler$Ship <- as.factor(SNEMA_springbiomass_species_compiler$Ship)
target_var <- "ChangeZAbs"
other_var <- colnames(SNEMA_springbiomass_species_compiler[3:32])
group_var <- "Ship"
SNEMA_spring_cor_results_allship <- group_reg_allship(SNEMA_springbiomass_species_compiler, target_var, other_var, group_var)
SNEMA_spring_cor <- SNEMA_spring_cor_results_allship %>% mutate(Scen = "SNEMA_spring_lowcorrelation_biomass")
correlation_filter_SNEMAspringbiomass <-SNEMA_spring_cor %>% filter(P_Value <= 0.05 & Coefficient > 0)
print(correlation_filter_SNEMAspringbiomass)

SNEMA_fallnumber_species_compiler <- as_tibble(SNEMA_fallnumber_species_compiler)
SNEMA_fallnumber_species_compiler$Ship <- as.factor(SNEMA_fallnumber_species_compiler$Ship)
target_var <- "ChangeZAbs"
other_var <- colnames(SNEMA_fallnumber_species_compiler[3:32])
group_var <- "Ship"
SNEMA_fall_cor_results_allship <- group_reg_allship(SNEMA_fallnumber_species_compiler, target_var, other_var, group_var)
SNEMA_fall_cor <- SNEMA_fall_cor_results_allship %>% mutate(Scen = "SNEMA_fall_lowcorrelation_number")
correlation_filter_SNEMAfallnumber <-SNEMA_fall_cor %>% filter(P_Value <= 0.05 & Coefficient > 0)
print(correlation_filter_SNEMAfallnumber)

SNEMA_fallbiomass_species_compiler <- as_tibble(SNEMA_fallbiomass_species_compiler)
SNEMA_fallbiomass_species_compiler$Ship <- as.factor(SNEMA_fallbiomass_species_compiler$Ship)
target_var <- "ChangeZAbs"
other_var <- colnames(SNEMA_fallbiomass_species_compiler[3:32])
group_var <- "Ship"
SNEMA_fall_cor_results_allship <- group_reg_allship(SNEMA_fallbiomass_species_compiler, target_var, other_var, group_var)
SNEMA_fall_cor <- SNEMA_fall_cor_results_allship %>% mutate(Scen = "SNEMA_fall_lowcorrelation_biomass")
correlation_filter_SNEMAfallbiomass <-SNEMA_fall_cor %>% filter(P_Value <= 0.05 & Coefficient > 0)
print(correlation_filter_SNEMAfallbiomass)

combinedd_mortality <- rbind(correlation_filter_GOMspringnumber, correlation_filter_GOMfallnumber, correlation_filter_GOMspringbiomass, correlation_filter_GOMfallbiomass, correlation_filter_SNEMAspringnumber, correlation_filter_SNEMAfallnumber, correlation_filter_SNEMAspringbiomass, correlation_filter_SNEMAfallbiomass)
combinedd_mortality
```

#Plots showing numbers of each species - catch rate
```{r}
#total number of each species

total_number <- combinedd %>% 
  mutate(species = stringr::str_remove(Other_Var, "_(wt|num)$"),
         unit = stringr::str_extract(Other_Var, "_(wt|num)$")) %>% 
  group_by(species) %>%
  summarise(count = n())

#without predator or competitor designation
total_number_plot <- data.frame(
  NameofSpecies=total_number$species ,  
  NumberofSpecies= total_number$count)

#with predator or competitor designation for removing those points 
total_number_plot <- data.frame(
  NameofSpecies=total_number$species ,  
  NumberofSpecies= total_number$count,
  PredorComp = c("Predator", "Competitor", "Competitor", "Competitor", "Both", "Competitor") #this is still hard coded; find some way to automate this later
  )

#without predator or competitor designation 
ggplot(total_number_plot, aes(x=NameofSpecies, y=NumberofSpecies)) + 
  geom_bar(stat = "identity") +
  ggtitle("Species With Significant Negative Relationships (LM) with Winter Flounder Catch Rates") +
  ylab("Number of instances where p<0.05") +
  xlab("species name")+
   coord_cartesian(ylim = c(0, 5)) +
  theme_classic() +
    theme(axis.title.x = element_blank(), axis.text.x = element_text(size=9, angle = 45, hjust = 1)) +
  theme(plot.title = element_text(size = 10))

#with predator or competitor designation 
ggplot(total_number_plot, aes(x=NameofSpecies, y=NumberofSpecies, fill = PredorComp)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Competitor" = "#154360", "Predator" = "gray", "Both" = "#FFC300")) + 
  ggtitle("Species With Significant Negative Relationships (LM) with Winter Flounder Catch Rates") +
  ylab("Number of instances where p<0.05") +
  xlab("species name")+
   coord_cartesian(ylim = c(0, 8)) +
  theme_classic() +
    theme(axis.title.x = element_blank(), axis.text.x = element_text(size=9, angle = 45, hjust = 1)) +
  theme(plot.title = element_text(size = 10)) + 
  theme(legend.position = "none") + 
  guides(fill=guide_legend(title="Predator or Competitor"))

#by abundance or biomass

#abundance
total_number_num <- combinedd %>% 
  filter(grepl("_num", Other_Var)) %>%
  mutate(species = stringr::str_remove(Other_Var, "_(wt|num)$"),
         unit = stringr::str_extract(Other_Var, "_(wt|num)$")) %>% 
  group_by(species) %>%
  summarise(count = n())

#without predator or competitor designation
total_number_plot_num <- data.frame(
  NameofSpecies=total_number_num$species ,  
  NumberofSpecies= total_number_num$count)

#with predator or competitor designation for removing those points 
total_number_plot_num <- data.frame(
  NameofSpecies=total_number_num$species ,  
  NumberofSpecies= total_number_num$count,
  PredorComp = c("Predator", "Competitor") #hard coded
  )

#without predator or competitor designation 
ggplot(total_number_plot_num, aes(x=NameofSpecies, y=NumberofSpecies)) + 
  geom_bar(stat = "identity") +
  ggtitle("Species With Significant Negative Relationships (LM) with Winter Flounder Abundance Catch Rates") +
  ylab("Number of instances where p<0.05") +
  xlab("species name")+
   coord_cartesian(ylim = c(0, 5)) +
  theme_classic() +
    theme(axis.title.x = element_blank(), axis.text.x = element_text(size=9, angle = 45, hjust = 1)) +
  theme(plot.title = element_text(size = 10))

#with predator or competitor designation 
ggplot(total_number_plot_num, aes(x=NameofSpecies, y=NumberofSpecies, fill = PredorComp)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Competitor" = "#154360", "Predator" = "gray", "Both" = "#FFC300")) + 
  ggtitle("Species With Significant Negative Relationships (LM) with Winter Flounder Abundance Catch Rates") +
  ylab("Number of instances where p<0.05") +
  xlab("species name")+
   coord_cartesian(ylim = c(0, 8)) +
  theme_classic() +
    theme(axis.title.x = element_blank(), axis.text.x = element_text(size=9, angle = 45, hjust = 1)) +
  theme(plot.title = element_text(size = 10)) + 
  theme(legend.position = "none") + 
  guides(fill=guide_legend(title="Predator or Competitor"))

#biomass
total_number_wt <- combinedd %>% 
  filter(grepl("_wt", Other_Var)) %>%
  mutate(species = stringr::str_remove(Other_Var, "_(wt|num)$"),
         unit = stringr::str_extract(Other_Var, "_(wt|num)$")) %>% 
  group_by(species) %>%
  summarise(count = n())

#without predator or competitor designation
total_number_plot_wt <- data.frame(
  NameofSpecies=total_number_wt$species ,  
  NumberofSpecies= total_number_wt$count)

total_number_plot_wt <- data.frame(
  NameofSpecies=total_number_wt$species ,  
  NumberofSpecies= total_number_wt$count,
  PredorComp = c("Predator", "Competitor", "Competitor", "Competitor", "Both", "Competitor") #this is still hard coded; find some way to automate this later
  )

#without predator or competitor designation 
ggplot(total_number_plot_wt, aes(x=NameofSpecies, y=NumberofSpecies)) + 
  geom_bar(stat = "identity") +
  ggtitle("Species With Significant Negative Relationships (LM) with Winter Flounder Biomass Catch Rates") +
  ylab("Number of instances where p<0.05") +
  xlab("species name")+
   coord_cartesian(ylim = c(0, 5)) +
  theme_classic() +
    theme(axis.title.x = element_blank(), axis.text.x = element_text(size=9, angle = 45, hjust = 1)) +
  theme(plot.title = element_text(size = 10))

#with predator or competitor designation 
ggplot(total_number_plot_wt, aes(x=NameofSpecies, y=NumberofSpecies, fill = PredorComp)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Competitor" = "#154360", "Predator" = "gray", "Both" = "#FFC300")) + 
  ggtitle("Species With Significant Negative Relationships (LM) with Winter Flounder Biomass Catch Rates") +
  ylab("Number of instances where p<0.05") +
  xlab("species name")+
   coord_cartesian(ylim = c(0, 5)) +
  theme_classic() +
    theme(axis.title.x = element_blank(), axis.text.x = element_text(size=9, angle = 45, hjust = 1)) +
  theme(plot.title = element_text(size = 10)) + 
  theme(legend.position = "none") + 
  guides(fill=guide_legend(title="Predator or Competitor"))
```

#Plots showing numbers of each species - mortality rate
```{r}
#total number of each species

total_number <- combinedd_mortality %>% 
  mutate(species = stringr::str_remove(Other_Var, "_(wt|num)$"),
         unit = stringr::str_extract(Other_Var, "_(wt|num)$")) %>% 
  group_by(species) %>%
  summarise(count = n())

#without predator or competitor designation
total_number_plot <- data.frame(
  NameofSpecies=total_number$species ,  
  NumberofSpecies= total_number$count)

#with predator or competitor designation for removing those points <- p value of predator/competitor species
total_number_plot <- data.frame(
  NameofSpecies=total_number$species ,  
  NumberofSpecies= total_number$count,
  PredorComp = c("Both", "Competitor", "Predator", "Both", "Competitor", "Both", "Both", "Predator", "Both", "Competitor", "Both") #this is still hard coded; find some way to automate this later
  )

#with predator or competitor designation for removing those points <- overall model p value 
#total_number_plot <- data.frame(
  #NameofSpecies=total_number$species ,  
  #NumberofSpecies= total_number$count,
  #PredorComp = c("Both", "Predator", "Both", "Competitor", "Both", "Both", "Competitor") #this is still hard coded; find some way to automate this later
  #)

#without predator or competitor designation 
ggplot(total_number_plot, aes(x=NameofSpecies, y=NumberofSpecies)) + 
  geom_bar(stat = "identity") +
  ggtitle("Species With Significant Positive Relationships (LM) with Winter Flounder Natural Mortality") +
  ylab("Number of instances where p<0.05") +
  xlab("species name")+
   coord_cartesian(ylim = c(0, 8)) + 
  theme_classic() +
    theme(axis.title.x = element_blank(), axis.text.x = element_text(size=9, angle = 45, hjust = 1)) +
  theme(plot.title = element_text(size = 10))

#with predator or competitor designation 
ggplot(total_number_plot, aes(x=NameofSpecies, y=NumberofSpecies, fill = PredorComp)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Competitor" = "#154360", "Predator" = "gray", "Both" = "#FFC300")) + 
  ggtitle("Species With Significant Positive Relationships (LM) with Winter Flounder Natural Mortality") +
  ylab("Number of instances where p<0.05") +
  xlab("species name")+
   coord_cartesian(ylim = c(0, 8)) +
  theme_classic() +
    theme(axis.title.x = element_blank(), axis.text.x = element_text(size=9, angle = 45, hjust = 1)) +
  theme(plot.title = element_text(size = 10)) + 
  theme(legend.position = "none") + 
  guides(fill=guide_legend(title="Predator or Competitor"))

#by abundance or biomass

#abundance
total_number_num <- combinedd_mortality %>% 
  filter(grepl("_num", Other_Var)) %>%
  mutate(species = stringr::str_remove(Other_Var, "_(wt|num)$"),
         unit = stringr::str_extract(Other_Var, "_(wt|num)$")) %>% 
  group_by(species) %>%
  summarise(count = n())

#without predator or competitor designation
total_number_plot_num <- data.frame(
  NameofSpecies=total_number_num$species ,  
  NumberofSpecies= total_number_num$count)

#with predator or competitor designation for removing those points 
total_number_plot_num <- data.frame(
  NameofSpecies=total_number_num$species ,  
  NumberofSpecies= total_number_num$count,
  PredorComp = c("Predator", "Both", "Competitor", "Both", "Both", "Competitor") #this is still hard coded; find some way to automate this later
  )

#without predator or competitor designation 
ggplot(total_number_plot_num, aes(x=NameofSpecies, y=NumberofSpecies)) + 
  geom_bar(stat = "identity") +
  ggtitle("Species With Significant Negative Relationships (LM) with Winter Flounder Mortality Rates") +
  ylab("Number of instances where p<0.05") +
  xlab("species name")+
   coord_cartesian(ylim = c(0, 5)) +
  theme_classic() +
    theme(axis.title.x = element_blank(), axis.text.x = element_text(size=9, angle = 45, hjust = 1)) +
  theme(plot.title = element_text(size = 10))

#with predator or competitor designation 
ggplot(total_number_plot_num, aes(x=NameofSpecies, y=NumberofSpecies, fill = PredorComp)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Competitor" = "#154360", "Predator" = "gray", "Both" = "#FFC300")) + 
  ggtitle("Species With Significant Negative Relationships (LM) with Winter Flounder Mortality Rates") +
  ylab("Number of instances where p<0.05") +
  xlab("species name")+
   coord_cartesian(ylim = c(0, 8)) +
  theme_classic() +
    theme(axis.title.x = element_blank(), axis.text.x = element_text(size=9, angle = 45, hjust = 1)) +
  theme(plot.title = element_text(size = 10)) + 
  theme(legend.position = "none") + 
  guides(fill=guide_legend(title="Predator or Competitor"))

#biomass
total_number_wt <- combinedd_mortality %>% 
  filter(grepl("_wt", Other_Var)) %>%
  mutate(species = stringr::str_remove(Other_Var, "_(wt|num)$"),
         unit = stringr::str_extract(Other_Var, "_(wt|num)$")) %>% 
  group_by(species) %>%
  summarise(count = n())

#without predator or competitor designation
total_number_plot_wt <- data.frame(
  NameofSpecies=total_number_wt$species ,  
  NumberofSpecies= total_number_wt$count)

total_number_plot_wt <- data.frame(
  NameofSpecies=total_number_wt$species ,  
  NumberofSpecies= total_number_wt$count,
  PredorComp = c("Both", "Competitor", "Predator", "Both", "Competitor", "Both", "Both", "Predator", "Both", "Competitor", "Both") #this is still hard coded; find some way to automate this later
  )

#without predator or competitor designation 
ggplot(total_number_plot_wt, aes(x=NameofSpecies, y=NumberofSpecies)) + 
  geom_bar(stat = "identity") +
  ggtitle("Species With Significant Negative Relationships (LM) with Winter Flounder Mortality Rates") +
  ylab("Number of instances where p<0.05") +
  xlab("species name")+
   coord_cartesian(ylim = c(0, 5)) +
  theme_classic() +
    theme(axis.title.x = element_blank(), axis.text.x = element_text(size=9, angle = 45, hjust = 1)) +
  theme(plot.title = element_text(size = 10))

#with predator or competitor designation 
ggplot(total_number_plot_wt, aes(x=NameofSpecies, y=NumberofSpecies, fill = PredorComp)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Competitor" = "#154360", "Predator" = "gray", "Both" = "#FFC300")) + 
  ggtitle("Species With Significant Negative Relationships (LM) with Winter Flounder Mortality Rates") +
  ylab("Number of instances where p<0.05") +
  xlab("species name")+
   coord_cartesian(ylim = c(0, 5)) +
  theme_classic() +
    theme(axis.title.x = element_blank(), axis.text.x = element_text(size=9, angle = 45, hjust = 1)) +
  theme(plot.title = element_text(size = 10)) + 
  theme(legend.position = "none") + 
  guides(fill=guide_legend(title="Predator or Competitor"))
```